---
title: "Analytic Graphics"
author: "han"
date: "2022-08-26"
output: html_document
---

# Analytic Graphics
Principles of analytic graphics

1. show comparisons

2. show causality, mechanism, explanation, systematic structure

3. show multivariate data

4. integration of evidence

5. describe and document the evidence with appropriate labels,scales, sources, etc

6. content is king

## Exploratory graphs
```{r}
with(airquality[!is.na(airquality$Ozone),], {
hist(airquality$Ozone,col="green",breaks = 20)
rug(Ozone)
    })

with(airquality,{
hist(Wind,col="green",breaks = 20)
abline(v=12,lwd=2)
abline(v=median(Wind),col="magenta",lwd=4)
})

with(airquality,{
  boxplot(airquality$Wind,color="blue")
  abline(h=10)
  
})


```
multiple variables
```{r}
boxplot(uptake ~ Treatment,data=CO2,col="red")

par(mfrow=c(2,1),mar=c(4,4,2,1))
hist(subset(CO2,Treatment=="nonchilled")$uptake,col="green")
hist(subset(CO2,Treatment=="chilled")$uptake,col="green")

par(mfrow=c(1,1))
plot(CO2$conc,CO2$uptake,col=CO2$Treatment)
abline(h=30,lwd=2,lty=2)

par(mfrow=c(1,2),mar=c(5,4,2,1))
with(subset(CO2,Treatment=="nonchilled"),{plot(conc,uptake,main="nonchilled")})
with(subset(CO2,Treatment=="chilled"),{plot(conc,uptake,main="chilled")})
```
## Graphics Devices
?Devices
windows()

1.Explicitly launch a graphics device

2. Call a plotting function to make a plot (Note: if you are using a file device, no plot will appear on the screen)

3.Annotate plot if necessary

4. Explicitly close graphics device with dev.off() (this is very important!)

```{r}
pdf(file = "myplot.pdf")
with(faithful,plot(eruptions,waiting))
title(main="old Faithful Geyser data")
dev.off()
```
### Graphics File Devices

There are two basic types of file devices:vector and bitmap devices 

Vector formats:

pdf:useful for line-type graphics, resizes well,usually portable, not efficient if a plot has many objects/points

svg:XML-based scalable vector graphics; supports animation and interactivity,potentially useful for web-based plots

win.metafile:Windows metafile format (only on Windows)

postscript:older format,also resizes well,usually portable,can be used to create encapsulated postscript files; Windows systems often don't have a postscript viewer

Bitmap formats
png: bitmapped format,good for line drawings or images with solid colors, uses lossless compression (like the old GIF format), most web browsers can read this format natively,good for plotting many many many points, does not resize well

jpeg:good for photographs or natural scenes,uses lossy compression,good for plotting many many many points, does not resize well, can be read byalmost any computer and any web browser,not great for line drawings

tiff:Creates bitmap files in the TIFF format; supports lossless compression

bmp:a native Windows bitmapped format

### Multiple Open Graphics Devices

It is possible to open multiple graphics devices(screen,file, or both), for example when viewing multiple plots at once

Plotting can only occur on one graphics device at a time

The currently active graphics device can be found by calling dev.cur()

Every open graphics device is assigned an integer ≥ 2.

You can change the active graphics device with dev.set(<integer>)where <integer> is the number associated with the graphics device you want to switch to


```{r}
dev.cur()
dev.set(1)
plot(x,y)
dev.set(3)
plot(y,z)
```

### Copying Plots

Copying a plot to another device can be useful because some plots require a lot of code and it can be a pain to type all that in again for a different device.

dev.copy:copy a plot from one device to another 

dev.copy2pdf:specifically copy a plot to a PDF file

NOTE: Copying a plot is not an exact operation, so the result may not be identical to the original. 
```{r}
library(datasets)

with(faithful, plot(eruptions, waiting)) ## Create plot on screen device
title(main ="0ld Faithful Geyser data") ## Add a main title

dev.copy(png, file = "geyserplot.png") ## Copy my plot to a PNG file 
dev.off() ## Don't forget to close the PNG device!

```

## Base Ploting System
?par
```{r}
library(datasets)
par(mfrow=c(1,1))
data(cars)
with(cars,plot(speed,dist))

hist(airquality$Ozone)
with(airquality,plot(Wind,Ozone))
boxplot(Ozone~Month,airquality)
airquality<-transform(airquality,Month=factor(Month))
boxplot(Ozone~Month,airquality)

```
pch:the plotting symbol (default is open circle)

lty:the line type (default is solid line),can be dashed, dotted,etc. 

lwd:the line width,specified as an integer multiple

col:the plotting color, specified as a number, string,or hex code;the colors() function gives you a vector of colors by name

xlab:character string for the x-axis label

ylab:character string for the y-axis label

The par() function is used to specify global graphics parameters that affect all plots in an R session. 
These parameters can be overridden when specified as arguments to specific plotting functions.

las:the orientation of the axis labels on the plot 

bg:the background color

mar:the margin size

oma:the outer margin size (default is 0 for all sides)

mfrow:number of plots per row, column (plots are filled row-wise)

mfcol: number of plots per row, column (plots are filled column-wise) 

Default values for global graphics parameters
```{r}
par("lty")
par("col")
par("pch")
par("bg")
par("mar") #bottom,left,top,right
par("mfrow")
```

plot:make a scatterplot, or other type of plot depending on the class of the object being plotted 

lines:add lines to a plot,given a vector x values and a corresponding vector of y values (or a 2- column matrix);this function just connects the dots

points:add points to a plot

text: add text labels to a plot using specified x,y coordinates

title:add annotations to x,y axis labels,title, subtitle,outer margin 

mtext:add arbitrary text to the margins (inner or outer) of the plot 

axis:adding axis ticks/labels

```{r}
with(airquality,plot(Wind,Ozone,main="Ozone and Wind in New York City"))
with(subset(airquality,Month == 5),points(Wind,Ozone,col="blue"))

with(airquality,plot(Wind,Ozone,main="Ozone and Wind in New York City",type="n"))
with(subset(airquality,Month == 5),points(Wind,Ozone,col="blue"))
with(subset(airquality,Month != 5),points(Wind,Ozone,col="red"))
legend("topright",pch=1,col=c("blue","red"),legend = c("May","Other Months"))

with(airquality,plot(Wind,Ozone,main="Ozone and Wind in New York City",pch=20))
model<-lm(Ozone~Wind,airquality)
abline(model,lwd=2)

par(mfrow=c(1,2))
with(airquality,{
  plot(Wind,Ozone,main="Ozone and Wind")
  plot(Solar.R,Ozone,main="Ozone and Solar Radiation")
  })

par(mfrow=c(1,3),mar=c(4,4,2,1),oma=c(0,0,2,0))
with(airquality,{
  plot(Wind,Ozone,main="Ozone and Wind")
  plot(Solar.R,Ozone,main="Ozone and Solar Radiation")
  plot(Temp,Ozone,main="Ozone and Temperature")
  mtext("Ozone and Weather in New York City", outer=T)
  })

par(mfrow=c(1,1),mar=c(4,4,2,2),oma=c(0,0,0,0))
x<-rnorm(100)
hist(x,n=15)
y<-rnorm(100)
plot(x,y)
z<-rnorm(100)
plot(x,z)

plot(x,y,pch=24,col="blue",bg="red")
example(points)

plot(x,y,pch=20)
title("Scatterplot")
text(-2,-2,"Label")
legend("topleft",legend="Data",pch=20)
fit<-lm(y~x)
abline(fit)
abline(fit,lwd=3,col="blue")
plot(x,y,xlab="Weight",ylab="Height",main="Scatterplot",pch=20)
legend("topright",legend="Data",pch=20)
fit<-lm(y~x)
abline(fit,lwd=3,col="red")
z<-rpois(100,2)
par("mar")
par(mar=c(2,2,1,1))
par(mfrow=c(2,1))
plot(x,y,pch=20)
plot(x,z,pch=19)

par(mfrow=c(1,2))
par(mar=c(4,4,2,2))
plot(x,y,pch=20)
plot(x,z,pch=19)

par(mfrow=c(2,2))
plot(x,y)
plot(x,z)
plot(y,z)
plot(y,x)

par(mfcol=c(2,2)) #different order
plot(x,y)
plot(x,z)
plot(y,z)
plot(y,x)

par(mfrow=c(1,1))
x<-rnorm(100)
y<-x+rnorm(100)
g<-gl(2,50,labels=c("Male","Female"))
str(g)
plot(x,y)

plot(x,y,type="n")
points(x[g=="Male"],y[g=="Male"],col="green")
points(x[g=="Female"],y[g=="Female"],col="blue")
```


## Lattice System
xyplot:this is the main function for creating scatterplots

bwplot:box-and-whiskers plots("boxplots")

histogram:histograms

stripplot: like a boxplot but with actual points 

dotplot:plot dots on "violin strings"

splom: scatterplot matrix; like pairs in base plotting system

levelplot, contourplot: for plotting "image" data


xyplot(y ~ x | g * g,data)

We use the formula notation here,hence the~.

On the left of the~is the y-axis variable, on the right is the x-axis variable 

levels: f and g are conditioning variables-they are optional

-the*indicates an interaction between two variables

The second argument is the data frame or list from which the variables in the formula should be looked up

-If no data frame or list is passed, then the parent frame is used. 

If no other arguments are passed,there are defaults that can be used. 
```{r}
library(lattice)
library(datasets)
state<-data.frame(state.x77,region=state.region)
xyplot(Life.Exp~Income|region,data=state,layout=c(4,1))

xyplot(Ozone ~ Wind, data=airquality)

airquality<-transform(airquality,Month=factor(Month))
xyplot(Ozone~Wind|Month,data=airquality,layout=c(5,1))
```
Lattice graphics functions return an object of class trellis

The print methods for lattice functions actually do the work of plotting the data on the graphics device.

Lattice functions return "plot objects" that can, in principle, be stored (but it's usually better to just save the code+data).

On the command line,trellis objects are auto-printed so that it appears the function is plotting the data

Lattice Panel Functions

```{r}
set.seed(10)
x<-rnorm(100)
f<-rep(0:1,each=50)
y<-x+f -f*x +rnorm(100,sd=0.5)
g<-factor(f,labels = c("Group 1","Group 2"))
xyplot(y~x |g,layout=c(2,1))

#custom panel function

xyplot(y ~ x |f,panel = function(x,y,...){
  panel.xyplot(x,y,...)
  panel.abline(h=median(y),lty=2)
  panel.lmline(x,y,col=2)
})
```
The lattice system is ideal for creating conditioning plots where you examine the same kind of plot under many different conditions

Panel functions can be specified/customized to modify what is plotted in each of the plot panels

## ggplot2 System
qplot

factors are important for indicating subsets of data; they should be labeled

```{r}
library(ggplot2)
data(mpg)
qplot(displ,hwy,data=mpg,color=drv)
qplot(displ,hwy,data=mpg,geom=c("point","smooth"))
qplot(hwy,data=mpg,fill=drv)
qplot(displ,hwy,data=mpg,facets=.~drv)

qplot(hwy,data=mpg,facets=drv~.,binwidth=2)
```

A data frame

aesthetic mappings: how data are mapped to color, size 

geoms:geometric objects like points, lines, shapes. 

facets: for conditional plots.

stats:statistical transformations like binning, quantiles, smoothing.

scales: what scale an aesthetic map uses (example: male = red, female = blue).

coordinate system 

```{r}
library(ggplot2)
g<-ggplot(mpg,aes(displ,cty))
summary(g)
g+geom_point()+facet_grid(.~drv)
```

Annotation

Labels: xlab(), ylab(), labs(), ggtitle()

Each of the “geom" functions has options to modify 

For things that only make sense globally, use theme() 
- Example: theme(legend.position = "none")

Two standard appearance themes are included 
- theme_gray(): The default theme (gray background) 
- theme_bw(): More stark/plain
```{r}
g+geom_point(color="steelblue",size=4,alpha=1/2)

g+geom_point(aes(color=drv),size=4,alpha=1/2)+
  labs(title="mpg", x=expression("log "* pm[2.5]))+
  geom_smooth(size=4,linetype=3,method="lm",se=FALSE)+
  theme_bw(base_family = "Times")
```

Axis Limits

```{r}
testdat<-data.frame(x=1:100,y=rnorm(100))
testdat[50,2]<-100
plot(testdat$x,testdat$y,type="l",ylim = c(-3,3))

g<-ggplot(testdat,aes(x,y))
g+geom_line()
g+geom_line()+ylim(-3,3) #Outlier missing
g+geom_line()+coord_cartesian(ylim=c(-3,3)) #Outlier included
```

cut
```{r}
ctycutpoints<-quantile(mpg$cty,seq(0,1,length=4),na.rm=TRUE)
mpg$ctycut<-cut(mpg$cty,ctycutpoints)
levels(mpg$ctycut)

g<-ggplot(mpg,aes(displ,drv))
g+geom_point(alpha=1/3)+
  geom_smooth(method="lm",se="F",color="steelblue")+
  theme_bw(base_family = "Avenir",base_size=20)+
  facet_wrap(drv~ctycut,ncol=4)+
  labs(title="mpg", x=expression("log "* pm[2.5]))
```

## Plotting and Color in R
heat.colors();topo.colors()

The function colors ( ) lists the names of colors you can use in any plotting function 
```{r}
x<-rnorm(10)
y<-rnorm(10)
par(mfrow=c(1,1))
plot(x,y,col=1)
y<-rnorm(10)
points(x,y,col=2)
y<-rnorm(10)
points(x,y,col=3)
image(volcano, col = hcl.colors(11, "purples", rev = TRUE))
filled.contour(volcano, nlevels = 10,
               color.palette = function(n, ...)
                   hcl.colors(n, "purples", rev = TRUE, ...))

colors()

colorRamp(c("red", "green"))( (0:4)/4 ) ## (x) , x in [0,1]
colorRampPalette(c("blue", "red"))( 4 ) ## (n)
## a ramp in opacity of blue values
colorRampPalette(c(rgb(0,0,1,1), rgb(0,0,1,0)), alpha = TRUE)(8)

pal<-colorRamp(c("red","blue"))
pal(0)
pal(1)
pal(0.5)
pal(seq(0,1,len=10))

pal<-colorRampPalette(c("red","yellow"))
pal(2)
pal(10)

library(RColorBrewer)
cols<-brewer.pal(3,"BuGn")
cols
pal<-colorRampPalette(cols)
par(mar=c(4,2,1,1))
image(volcano,col=pal(20))

x<-rnorm(10000)
y<-rnorm(10000)
smoothScatter(x,y) #2Dhistgram
```
There are 3 types of palettes

-Sequential(YlOrRd)

-Diverging (Spectral)

-Qualitative (Set3)

colorspace package

rgb alpha
```{r}
plot(x,y,pch=19)
plot(x,y,pch=19,col=rgb(0,0,0,0.02))
```
## Clustering
### Hierarchical Clustering
·An agglomerative approach

-Find closest two things 

-Put them together 

-Find next closest 

Requires

-A defined distance 

-A merging approach 

Produces

-A tree showing how close things are to each other

·Most important step

·Garbage in->garbage out 

Distance or similarity

-Continuous-euclidean distance 

-Continuous-correlation similarity 

-Binary-manhattan distance

·Pick a distance/similarity that makes sense for your problem 

```{r}
set.seed(1234)
par(mar=c(0,0,0,0))
x<-rnorm(12,mean=rep(1:3,each=4),sd=0.2)
y<-rnorm(12,mean = rep(c(1,2,1),each=4),sd=0.2)
plot(x,y,col="blue",pch=19,cex=2)
text(x+0.05,y+0.05,labels=as.character(1:12))

dataFrame<-data.frame(x,y)
distxy<-dist(dataFrame)
hClustering<-hclust(distxy)
par(mar=c(4,4,2,2))
plot(hClustering)
```

```{r}
myplclust <-function(hclust,lab=hclust$labels,lab.col=rep(1,length(hclust$labels)), hang=0.1, ...){

## modifiction of plclust for plotting hclust objects *in colour*! Copyright 
##Eva KF Chan 2009 Arguments:hclust:hclust object lab:a character vector 
## of labels of the leaves of the tree lab.col: colour for the labels;
## NA=default device foreground colour hang: as in hclust & plclust Side 
## effect:A display of hierarchical cluster with coloured leaf labels.
  y<-rep(hclust$height,2)
  x<-as.numeric(hclust$merge) 
  y <- y[which(x<0)]
  x <- x[which(x<0)] 
  x<-abs(x)
  y<-y[order(x)]
  x<-x[order(x)]
  plot(hclust,labels=FALSE,hang=hang,...)
  text(x=x,y= y[hclust$order] -(max(hclust$height)*hang),labels = lab[hclust$order], col=lab.col[hclust$order],srt=90,adj=c(1,0.5),xpd=NA,...) 
}

dataFrame<-data.frame(x,y)
distxy<-dist(dataFrame)
hClustering<-hclust(distxy)
par(mar=c(4,4,2,2))
myplclust(hClustering,lab=rep(1:3,each=4),lab.col=rep(1:3,each=4))

set.seed(143)
dataMarix<-as.matrix(dataFrame)[sample(1:12),]
heatmap(dataMarix)
```
·Gives an idea of the relationships between variables/observations 

The picture may be unstable

But it is deterministic

Choosing where to cut isn't always obvious 

Should be primarily used for exploration

### K-means Clustering
A partioning approach

-Fix a number of clusters 

-Get"centroids" of each cluster 

-Assign things to closest centroid

-Reclaculate centroids

·Requires

-A defined distance metric 

-A number of clusters

-An initial guess as to cluster centroids

·Produces

-Final estimate of cluster centroids 

-An assignment of each point to clusters 

```{r}
set.seed(1234)
par(mar=c(0,0,0,0))
x<-rnorm(12,mean=rep(1:3,each=4),sd=0.2)
y<-rnorm(12,mean = rep(c(1,2,1),each=4),sd=0.2)
plot(x,y,col="blue",pch=19,cex=2)
text(x+0.05,y+0.05,labels=as.character(1:12))

dataFrame<-data.frame(x,y)

kmeansObj<-kmeans(dataFrame,centers=3)
names(kmeansObj)
kmeansObj$cluster

par(mar=rep(0.2,4))
plot(x,y,col=kmeansObj$cluster,pch=19,cex=2)
points(kmeansObj$centers,col=1:3,pch=3,cex=3,lwd=3)


set.seed(143)
dataMarix<-as.matrix(dataFrame)[sample(1:12),]
kmeansObj2<-kmeans(dataMarix,centers=3)
par(mfrow=c(1,2),mar=c(2,4,0.1,0.1))
image(t(dataMarix)[,nrow(dataMarix):1],yaxt="n")
image(t(dataMarix)[,order(kmeansObj$cluster)],yaxt="n")

```
·K-means requires a number of clusters 

-Pick by eye/intuition

-Pick by cross validation/information theory,etc.

-Determiningthe number of clusters

·K-means is not deterministic 

-Different#of clusters

-Different number of iterations 

### Dimension Reduction

```{r}
set.seed(12345)
par(mfrow=c(1,1),mar=rep(4,4))
dataMarix<-matrix(rnorm(400),nrow=40)
image(1:10,1:40,t(dataMarix)[,nrow(dataMarix):1])
heatmap(dataMarix)

set.seed(678910)
for (i in 1:40) {
  coinFlip<-rbinom(1,size=1,prob=0.5)
  if(coinFlip){
    dataMarix[i,]<-dataMarix[i,]+rep(c(0,3),each=5)
  }
  }
par(mar=rep(0.2,4))
image(1:10,1:40,t(dataMarix)[,nrow(dataMarix):1])
heatmap(dataMarix)

hh<-hclust(dist(dataMarix))
dataMatrixOrdered<-dataMarix[hh$order,]
par(mfrow=c(1,3))
image(t(dataMatrixOrdered)[,nrow(dataMatrixOrdered):1])
plot(rowMeans(dataMatrixOrdered),40:1,xlab = "Row Mean",ylab = "Row",pch=19)
plot(rowMeans(dataMatrixOrdered),xlab = "Column",ylab = "Column Mean",pch=19)
```
You have multivariate variables X1,Xn soX1=(X11,...,X1m)

Find a new set of multivariate variables that are uncorrelated and explain as much variance as possible.

If you put all the variables together in one matrix, find the best matrix created with fewer variables (lower rank)that explains the original data.

The first goal is statistical and the second goal is data compression. 

SVD

If X is a matrix with each variable in a column and each observation in a row then the SVD is a"matrix decomposition"

X=UDV^T

where the columns of U are orthogonal (left singular vectors), the columns of V are orthogonal (right singular vectors) and D is a diagonal matrix (singular values).

PCA

The principal components are equal to the right singular values if you first scale (subtract the mean, divide by the standard deviation)the variables.
```{r}
svd1<-svd(scale(dataMatrixOrdered))
par(mfrow=c(1,3))
image(t(dataMatrixOrdered)[,nrow(dataMatrixOrdered):1])
plot(svd1$u[,1],40:1,xlab = "Row",ylab = "First left singular vector",pch=19)
plot(svd1$v[,1],xlab = "Column",ylab = "First right singular vector",pch=19)

par(mfrow=c(1,2))
plot(svd1$d,xlab="Column",ylab="Singular value",pch=19)
plot(svd1$d^2/sum(svd1$d^2),xlab="Column",ylab="Prop. of variance explained",pch=19)

# Relationship to principal components
svd1<-svd(scale(dataMatrixOrdered))
pca1<-prcomp(dataMatrixOrdered,scale=TRUE)
plot(pca1$rotation[,1],svd1$v[,1],pch=19,xlab="Principal Component 1",ylab="Right Singular Vector 1")
abline(c(0,1))

constantMatrix<-dataMatrixOrdered*0
for (i in 1:dim(dataMatrixOrdered)[1]) {
  constantMatrix[i,]<-rep(c(0,1),each=5)
}
svd1<-svd(constantMatrix)
par(mfrow=c(1,3))
image(t(constantMatrix)[,nrow(constantMatrix):1])
plot(svd1$d,xlab="Column",ylab="Singular value",pch=19)
plot(svd1$d^2/sum(svd1$d^2),xlab="Column",ylab="Prop. of variance explained",pch=19)

set.seed(678910)
for (i in 1:40) {
  coinFlip1<-rbinom(1,size=1,prob=0.5)
  coinFlip2<-rbinom(1,size=1,prob=0.5)
  if (coinFlip1){
    dataMarix[i,]<-dataMarix[i,]+rep(c(0,5),each=5)
  }
  if (coinFlip2){
    dataMarix[i,]<-dataMarix[i,]+rep(c(0,5),5)
  }
}
  hh<-hclust(dist(dataMarix))
  dataMatrixOrdered<-dataMarix[hh$order,]

svd2<-svd(scale(dataMatrixOrdered))
par(mfrow=c(1,3))
image(t(dataMatrixOrdered))
plot(rep(c(0,1),each=5),pch=19,xlab="Column",ylab="Pattern 1")
plot(rep(c(0,1),5),pch=19,xlab="Column",ylab="Pattern 2")

svd2<-svd(scale(dataMatrixOrdered))
par(mfrow=c(1,3))
image(t(dataMatrixOrdered))
plot(svd2$v[,1],pch=19,xlab="Column",ylab="First right singular vector")
plot(svd2$v[,2],pch=19,xlab="Column",ylab="Second right singular vector")


svd1<-svd(scale(dataMatrixOrdered))
par(mfrow=c(1,2))
plot(svd1$d,xlab="Column",ylab="Singular value",pch=19)
plot(svd1$d^2/sum(svd1$d^2),xlab="Column",ylab="Prop. of variance explained",pch=19)

```
scale函数是将一组数进行处理，默认情况下是将一组数的每个数都减去这组数的平均值后再除以这组数的均方根

Missing values

```{r}
dataMarix2<-dataMatrixOrdered
dataMarix2[sample(1:100,size=40,replace=FALSE)]<-NA
# svd1<-svd(scale(dataMarix2))

if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

BiocManager::install("impute")
library(impute)
dataMarix2<-dataMatrixOrdered
dataMarix2[sample(1:100,size=40,replace=FALSE)]<-NA
dataMarix2<-impute.knn(dataMarix2$data)
svd1<-svd(scale(dataMatrixOrdered))
svd2<-svd(scale(dataMarix2))
par(mfrow=c(1,2));plot(svd1$v[,1],pch=19);plot(svd2$v,pch=19)
# svd1<-svd(scale(dataMarix2))

```
face example
```{R}
load("data/face.rda")
par(mfrow=c(1,1))
image(t(faceData)[,nrow(faceData):1])
svd1<-svd(scale(faceData))
plot(svd1$d^2/sum(svd1$d^2),xlab="Column",ylab="Prop. of variance explained",pch=19)

approx1<-svd1$u[,1] %*% t(svd1$v[,1]) *svd1$d[1]

approx5<-svd1$u[,1:5] %*% diag(svd1$d[1:5]) %*% t(svd1$v[,1:5]) 
approx10<-svd1$u[,1:10] %*% diag(svd1$d[1:10]) %*% t(svd1$v[,1:10]) 
par(mfrow=c(1,4),mar=c(2,2,0,0))
image(t(approx1)[,nrow(approx1):1],main="(a)")
image(t(approx5)[,nrow(approx5):1],main="(b)")
image(t(approx10)[,nrow(approx10):1],main="(c)")
image(t(faceData)[,nrow(faceData):1],main="(d)")
```
### Case study
#### Human activity with smart phones
```{r}
load("data/samsungData.rda")
names(samsungData)[1:12]
table(samsungData$activity)

par(mfrow=c(1,2),mar=c(5,4,1,1))
samsungData<-transform(samsungData,activity=factor(activity))
sub1<-subset(samsungData, subject==1)
plot(sub1[,1],col=sub1$activity,ylab=names(sub1)[1])
plot(sub1[,2],col=sub1$activity,ylab=names(sub1)[2])
legend("bottomright",legend=unique(sub1$activity),col=unique(sub1$activity),pch=1)

source("data/mypclust.R")
distanceMatrix<-dist(sub1[,1:3])
hclustering<-hclust(distanceMatrix)
par(mfrow=c(1,1))
myplclust(hclustering, lab.col = unclass(sub1$activity))

par(mfrow=c(1,2))
plot(sub1[,10],pch=19,col=sub1$activity,ylab=names(sub1)[10])
plot(sub1[,11],pch=19,col=sub1$activity,ylab=names(sub1)[11])

distanceMatrix<-dist(sub1[,10:12])
hclustering<-hclust(distanceMatrix)
par(mfrow=c(1,1))
myplclust(hclustering, lab.col = unclass(sub1$activity))

svd1<-svd(scale(sub1[,-c(562,563)]))
par(mfrow=c(1,2))
plot(svd1$u[,1],col=sub1$activity,pch=19)
plot(svd1$u[,2],col=sub1$activity,pch=19)

par(mfrow=c(1,1))
plot(svd1$u[,2],pch=19)

maxContrib<-which.max(svd1$v[,2])
distanceMatrix<-dist(sub1[,c(10:12,maxContrib)])
hclustering<-hclust(distanceMatrix)
myplclust(hclustering,lab.col = unclass(sub1$activity))

names(samsungData)[maxContrib]

kClust<-kmeans(sub1[,-c(562,563)],centers = 6)
table(kClust$cluster,sub1$activity)

kClust<-kmeans(sub1[,-c(562,563)],centers = 6,nstart = 1)
table(kClust$cluster,sub1$activity)

kClust<-kmeans(sub1[,-c(562,563)],centers = 6,nstart = 100)
table(kClust$cluster,sub1$activity)

plot(kClust$centers[1,1:10],pch=19,ylab="Cluster Center",xlab="")

plot(kClust$centers[4,1:10],pch=19,ylab="Cluster Center",xlab="")
```
## Air Pollution Case Study
```{r}
pm0 <- read.table("data/RD_501_88101_1999-0.txt",comment.char = "#",header = FALSE,sep="|",na.strings = "")
dim(pm0)
head(pm0)
cnames<-readLines("data/RD_501_88101_1999-0.txt",1)
cnames
cnames<-strsplit(cnames,"|",fixed=TRUE)
cnames
names(pm0)<-make.names(cnames[[1]])
head(pm0)
x0<-pm0$Sample.Value
class(x0)
str(x0)
summary(x0)
mean(is.na(x0))

pm1 <- read.table("data/RD_501_88101_2012-0.txt",comment.char = "#",header = FALSE,sep="|",na.strings = "")
dim(pm1)
names(pm1)<-make.names(cnames[[1]])
head(pm1)
x1<-pm1$Sample.Value
str(x1)
summary(x1)
summary(x0)
mean(is.na(x1))

boxplot(x0,x1)
boxplot(log10(x0),log10(x1))

negative<-x1<0
str(negative)
sum(negative,na.rm=TRUE)
mean(negative,na.rm=TRUE)
dates<-pm1$Date
str(dates)
dates<-as.Date(as.character(dates),"%Y%m%d")
str(dates)
hist(dates,"month")
hist(dates[negative],"month")

site0<- unique(subset(pm0,State.Code ==36, c(County.Code, Site.ID)))
site1<- unique(subset(pm1,State.Code ==36, c(County.Code, Site.ID)))
head(site0)
site0<-paste(site0[,1],site0[,2],sep=".")
site1<-paste(site1[,1],site1[,2],sep=".")

str(site0)
str(site1)
both<-intersect(site0,site1)
both
pm0$county.site<-with(pm0,paste(County.Code,Site.ID,sep="."))
pm1$county.site<-with(pm1,paste(County.Code,Site.ID,sep="."))
cnt0<-subset(pm0,State.Code==36 & county.site %in% both)
cnt1<-subset(pm1,State.Code==36 & county.site %in% both)
head(cnt0)
sapply(split(cnt0,cnt0$county.site), nrow)
sapply(split(cnt1,cnt1$county.site), nrow)

pm1sub<-subset(pm1,State.Code==36 & County.Code ==63 & Site.ID==2008)
pm0sub<-subset(pm0,State.Code==36 & County.Code ==63 & Site.ID==2008)
dim(pm1sub)
dim(pm0sub)
dates1 <- as.Date(as.character(pm1sub$Date),"%Y%m%d")
x1sub<-pm1sub$Sample.Value
plot(dates1,x1sub)

dates0 <- as.Date(as.character(pm0sub$Date),"%Y%m%d")
x0sub<-pm0sub$Sample.Value
plot(dates0,x0sub)

par(mfrow=c(1,2),mar=c(4,4,2,1))
plot(dates0,x0sub,pch=20)
abline(h=median(x0sub,na.rm=T))
plot(dates1,x1sub,pch=20)
abline(h=median(x1sub,na.rm=T))

rng<-range(x0sub,x1sub,na.rm=T)
plot(dates0,x0sub,pch=20,ylim=rng)
abline(h=median(x0sub,na.rm=T))
plot(dates1,x1sub,pch=20,ylim=rng)
abline(h=median(x1sub,na.rm=T))

head(pm0)
mn0<-with(pm0,tapply(Sample.Value,State.Code,mean,na.rm=T))
str(mn0)
summary(mn0)
mn1<-with(pm1,tapply(Sample.Value,State.Code,mean,na.rm=T))
str(mn1)
summary(mn1)
d0<-data.frame(state=names(mn0),mean=mn0)
d1<-data.frame(state=names(mn1),mean=mn1)
mrg<-merge(d0,d1,by="state")
dim(mrg)
head(mrg)
par(mfrow=c(1,1))
with(mrg,plot(rep(1999,52),mrg[,2],xlim=c(1998,2013)))
with(mrg,points(rep(2012,52),mrg[,3]))
segments(rep(1999,52),mrg[,2],rep(2012,52),mrg[,3])
```
